Which groups are at highest risk for low uptake of flu vaccine?

### 2023-04-21 products/manuscript/Manuscript.qmd
Results, discussion, and conclusion updates

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(png)

@@

```{r, warning=FALSE, message=FALSE}
result_table1 <- readRDS(here("results", "table1_demographics.rds"))
result_table1
```

@@

@@

*Summarize what you did, what you found and what it means.*

[\*\*To be completed (PART 5)]{.underline}

@@

*Discuss what you perceive as strengths and limitations of your
analysis.*

[\*\*To be completed (PART 5)]{.underline}

@@

*What are the main take-home messages?*

[\*\*To be completed (PART 5)]{.underline}

@@

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

[\*\*To be completed (PART 5)]{.underline}

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
library(png)

@@

```{r, warning=FALSE, message=FALSE, echo=FALSE}
result_table1 <- readRDS(here("results", "table1_demographics.rds"))
result_table1
```

@@

The results of our analysis showed that age was significantly associated with flu vaccine uptake (OR = 1.031, 95% CI = 1.030-1.032, p < 0.001), with older individuals being more likely to have received the vaccine. The sex of the individual was also significantly associated with flu vaccine uptake (OR = 1.416, 95% CI = 1.384-1.448, p < 0.001), with females being more likely to have received the vaccine compared to males. Race was also significantly associated with flu vaccine uptake. Specifically, individuals who identified as black (OR = 0.790, 95% CI = 0.776-0.805, p < 0.001) were less likely to have received the vaccine compared to those who identified as white. Native Americans (OR = 1.284, 95% CI = 1.149-1.436, p = 0.0001) and Asians (OR = 1.141, 95% CI = 1.089-1.195, p < 0.001) were more likely to have received the vaccine than whites. Hispanic individuals (OR = 0.825, 95% CI = 0.806-0.845, p < 0.001) were less likely to have received the vaccine compared to non-Hispanic individuals. Education level was also significantly associated with flu vaccine uptake, with individuals who had higher levels of education being more likely to have received the vaccine. Specifically, individuals with some college education (OR = 1.172, 95% CI = 1.126-1.220, p < 0.001), a bachelor's degree (OR = 1.460, 95% CI = 1.402-1.520, p < 0.001), or a graduate degree (OR = 1.911, 95% CI = 1.818-2.007, p < 0.001) were more likely to have received the vaccine compared to those with no education. Individuals with family incomes of $50,000-$99,999 (OR = 1.140, 95% CI = 1.112-1.168, p < 0.001) and those with family incomes of $100,000 or more (OR = 1.417, 95% CI = 1.380-1.455, p < 0.001) were more likely to have received the vaccine compared to those with family incomes less than $50,000. In terms of health-related variables, individuals with asthma (OR = 1.398, 95% CI = 1.361-1.435, p < 0.001), cancer (OR = 1.322, 95% CI = 1.281-1.364, p < 0.001), and coronary heart disease (OR = 1.637, 95% CI = 1.560-1.717, p < 0.001) were more likely to have received the vaccine compared to those without these conditions. Finally, smoking status was significantly associated with flu vaccine uptake, with former smokers (OR = 1.059, 95% CI = 1.032-1.087, p = 0.0006) being more likely to receive the vaccine compared to nonsmoker, and current smokers (OR = 0.669, 95% CI = 0.643 - 0.696) being less likely to receive vaccine. Finally, former drinkers (OR = 1.059, 95% CI = 1.010-1.111, p = 0.0172) were more likely to receive the vaccine compared to nondrinkers, and this was statistically significant. However, current drinkers (OR = 0.978, 95% CI = 0.941 - 1.016) were less likely to receive vaccine compared to nondrinkers. Although this association was not statistically significant.

@@

The results of our analysis indicate that several demographic and health-related factors are significantly associated with flu vaccine uptake. Our findings suggest that age, sex, race, education level, family income, and health status are all important predictors of vaccine uptake. Specifically, older individuals, females, individuals with higher education levels, and those with higher family incomes were more likely to have received the vaccine. On the other hand, individuals who identified as black or Hispanic were less likely to have received the vaccine, while Native Americans and Asians were more likely to have received it. In terms of health-related variables, individuals with asthma, cancer, and coronary heart disease were more likely to have received the vaccine compared to those without these conditions. Finally, our findings also suggest that smoking and drinking status are associated with vaccine uptake, with former smokers and drinkers more likely to have received the vaccine compared to nonsmokers and nondrinkers, respectively. These results have important implications for public health efforts aimed at increasing vaccine uptake, particularly among underserved and marginalized populations.

@@

While caution should be exercised when interpreting the significance of associations in observational studies due to potential confounding and reverse causality, this analysis provides valuable insights into identifying groups at risk for low influenza vaccine uptake over time. Limited studies have explored the relationship between a comprehensive list of factors and influenza vaccine coverage, and the large study population in this research improves statistical power, generalizability, precision, and accuracy of the findings. The design of the NHIS survey also allows for representative sampling of households, reducing sampling bias. Additionally, the survey questions were tailored to the month of each interview, enabling more detailed assessment of the seasonality of influenza vaccine schedule in the United States.

In terms of limitations, as noted above, observational studies cannot demonstrate causality. The reason for not receiving an influenza vaccine is not an NHIS variable, which could have provided more insight. It is also important to note that the primary outcome of vaccination within the past 12 months was self-reported, and there was no way to verify using medical records, meaning that there is a risk of reporting bias due to recall, social desirability, or other factors. The reliability of self-report compared to health records for influenza vaccine is not known. Lastly, It is also unclear to what extent the COVID-19 pandemic and antivaccine movements have impacted vaccine behavior and hesitancy, which could be explored in future qualitative studies.

@@

In conclusion, this study highlights the various demographic and health-related factors that are significantly associated with flu vaccine uptake. Understanding these predictors can help inform public health efforts aimed at increasing vaccine uptake, especially among marginalized populations. The findings suggest that targeted interventions may be necessary to address disparities in vaccine uptake among different racial and ethnic groups, as well as those with lower education levels and family incomes. Additionally, our findings on the association between smoking and drinking status and vaccine uptake may suggest the need for tailored messaging to encourage vaccination among these populations. Overall, these results can help guide public health interventions to promote vaccine uptake and ultimately reduce the burden of flu-related illness and hospitalizations.

@@
